---
title: "Week-10-DataDive-GLMs"
author: "Raghuveer Venkatesh"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
## Initial Setup and loading the Data
# Loading library and reading the data set
library(tidyverse)
library(ggthemes)
library(ggrepel)
library(boot)
library(broom)
library(lindia)
library(tidyverse)
library(readr)
library(dplyr)
library(lubridate)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(glue)
library(scales)
library(car)
# Set working directory as the path to the data set
setwd("C:/Users/raghu/OneDrive/Documents/Statistics_with_R/Week 2 Data Dive")

# Read the data set
player_data <- read_csv("fifa_players.csv")
# Inspecting the column
spec(player_data)
```

# Introduction

In this project, we explore preferred foot (right vs. left) as a binary variable to see how it relates to player skills and physical attributes. Using logistic regression, we’ll interpret the effects of key attributes, build a confidence interval for one coefficient, and discuss the insights gained along with potential follow-up questions.

## Why Preferred foot ?

This column can be converted into binary value easily Left Foot = 0, Right Foot = 1. Preferred foot can significantly influence a player's performance in certain positions or play styles, which might correlate with overall ratings, specific skill metrics (like crossing or finishing), or market value.


# Logistic Regression Model for Preferred foot

```{r}

#Encoding the values as 0 and 1.

player_data <- player_data |>
  mutate(preferred_foot_binary = ifelse(preferred_foot == "Right", 1, 0))

# Select the explanatory variables
explanatory_vars <- c("agility", "ball_control", "crossing", "dribbling")


model <- glm(preferred_foot_binary ~ agility + ball_control + crossing + dribbling,
             data = player_data, family = binomial(link = "logit"))


summary(model)

exp(coef(model))
```

## Explaining the co-efficients

  Intercept: 1.997097 
  **Odds Ratio**: exp(1.9971) ≈ 7.37
  This suggests that at baseline, odds are a player is 7.37 times more likely to be right-footed than left-footed.
  
   Agility
   **Odds Ratio**: exp(0.0050) ≈ 1.005
   Each point increase in agility raises the odds of being right footed by 0.5%
   
   Ball Control: 0.0309
  **Odds Ratio**: exp(0.0309) ≈ 1.031
   For each additional point in ball control, the odds of being right-footed increase by approximately 3.1%, indicating that ball control is positively associated    with right-foot preference.

   Crossing: -0.0545
   **Odds Ratio**: exp(-0.0545) ≈ 0.947
   Each additional point in crossing skill slightly decreases the odds of being right-footed by about 5.3%, suggesting that players with higher crossing ability      may be more likely to prefer their left foot.
  
   Dribbling: -0.0021
   **Odds Ratio**: exp(-0.0021) ≈ 0.998
   With an odds ratio near 1, each point increase in dribbling has no meaningful impact on the odds of being right-footed, indicating that dribbling skill does not    significantly influence foot preference.

   
### P - Value


Ball control and Crossing have very small p-values < 0.05 indicating that they are having a significant effect on Preferred Foot. Whereas Agility and Dribbling have a p-value > 0.05 especially dribbling with a p-value of 0.5452 so they have an influence on whether a player is Right-Footed.

## Plotting the model using Agility

```{r}

player_data$predicted_prob <- predict(model, type = "response")

# Create a plot
ggplot(player_data, aes(x = agility, y = predicted_prob)) +
  geom_point(aes(color = preferred_foot), alpha = 0.5) +  # Original data points
  geom_line(stat = "smooth", method = "glm", method.args = list(family = "binomial"), color = "blue") +  # Logistic regression curve
  labs(title = "Logistic Regression: Predicted Probability of Being Right-Footed",
       x = "Agility Score",
       y = "Predicted Probability of Being Right-Footed") +
  scale_color_manual(values = c("lightblue", "lightcoral"), 
                     name = "Preferred Foot", 
                     labels = c("Left-Footed", "Right-Footed")) +
  theme_minimal()

```
### Interpretation

As Agility score increases it's most probable that the player is Left footed.


# Confidence Interval and Standard error for Agility


```{r}
ggplot(player_data, aes(x = preferred_foot, y = agility, fill = preferred_foot)) +
  geom_boxplot() +
  labs(title = "Distribution of Agility by Foot Preference",
       x = "Foot Preference",
       y = "Agility Score") +
  theme_minimal() +
  scale_fill_manual(values = c("lightblue", "lightcoral"))

```

```{r}

coef_agility <- coef(model)["agility"]
se_agility <- summary(model)$coefficients["agility", "Std. Error"]


# Calculate the 95% Confidence Interval for agility coefficient
ci_lower <- coef_agility - 1.96 * se_agility
ci_upper <- coef_agility + 1.96 * se_agility

# Exponentiation of the CI to interpret it as an odds ratio range
ci_odds_ratio <- exp(c(ci_lower, ci_upper))

# Print results
cat("Odds Ratio for Agility:", exp(coef_agility), "\n")
cat("95% Confidence Interval for Odds Ratio of Agility:", ci_odds_ratio, "\n")

```
### Interpretation

For each one-point increase in agility, the odds of a player being right-footed increase by approximately 0.5%.


CI ranges from 1.0011 to 1.0090 meaning range of values that we are 95% confident contains the true odds ratio of agility in the population from which the sample was drawn.


Since it's positive, that means agility is positively associated with being right-footed meaning as the player has higher agility they are more likely to be Right footed.

## Further questions

How does these findings vary across different nationalities ?. 

Does it have the same impact on players with high rating as well as players with low rating ?




### THE END